---
title: "MED263 R Tutorial"
author: "Jihoon Kim (j5kim@ucsd.edu)"
date: "1/12/2017"
output: html_document
fontsize: 10pt
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

### R  

R is 'GNU S', a freely available language and environment for statistical computing and graphics which provides a wide variety of statistical and graphical techniques

- linear and nonlinear modeling
- statistial tests
- time series analysis
- classification
- clustering
- genomics


Download and install R (precompiled binary distribution) from UCLA <http://cran.stat.ucla.edu>, the nearest R mirror site <http://cran.r-project.org/mirrors.html> 


### R Studio 

RStudio is a set of integrated development environment (IDE) designed to help you be more prudictive with R. It includes 

- console
- syntax-highlighting editor that supports direct code execution
- plotting
- history
- debugging 
- workspace management. 

Download and install RStudio from <https://www.rstudio.com/products/rstudio/download>


### Calculate some numbers

```{r, echo = TRUE }
1 + 2
2^4
log(16, 2)
log(1000, 3)
choose(5, 2)
```


### Data types 

##### Number
```{r, echo = TRUE }
a <- 3
a
typeof( a )
b <- 5
a + b
```


##### Vector of numbers
```{r, echo = TRUE }
cholesterol <- c(177, 193, 195, 209, 226)
typeof( cholesterol )
length( cholesterol )
mean( cholesterol )
max( cholesterol )
```

##### Vector of characters
```{r, echo = TRUE }
genotype <- c("AA", "AG", "GG")
typeof( genotype )
```

##### Factors
```{r, echo = TRUE }
alleles <- c( "AG", "GG", "AA", "AA", "AA", "AA", "AA", "AG", "AG", "GG", "GG", "GG", "GG", "AA")
typeof( alleles )
alleles 
alleles <- factor( alleles )
typeof( alleles )
alleles
table( alleles )
```


##### Data Frames 
```{r, echo = TRUE }
```